N-dimensional Tic Tac Toe, and Adventure in Modules

### 2017-05-12 paper/paper.tex
editing for style

  As we touched on before - the notion of a game is generalized in the
  \texttt{game.sml} file. This defines the signatures which make up the pattern
  of a game and a functor to run a game.
  The general pattern for a game under our model consists of three purposely
  isolated pieces.
  These are \textit{State}, \textit{Actions}, and \textit{Agents} - which are
  each given their own signature in \texttt{game.sml}.
  These are then wrapped together with a functor to execute a game.
  The idea is that any game consists of a state, a set of actions on that state,
  and an agents that take actions for a state. A specific implementation of a
  game would use the relationship between these three general pieces to define
  a runnable game with its own modules.

@@

  In \texttt{tictactoe.sml} we have two signatures - one that extends state
  (the \texttt{STATE} signature from \texttt{game.sml}) for \textit{Tic-tac-toe}
  called \texttt{TTTSTATE} and another that does the same with action
  (the \texttt{ACTION} signature from \texttt{game.sml}) called
  \texttt{TTTACTION}.
  We don't extend the \texttt{AGENT} signature specifically for \textit{Tic-tac-toe}.

  There are many module structures instantiated in \texttt{tictactoe.sml}
  for the various kinds of state, actions and agents that we eventually want to
  use - such as \texttt{TttState} and \texttt{TttAction} for instance - but each will
  be used to implement the next kind of module.
  These module structures are bound to functors like \texttt{TttStateFn} which
  takes a module implementing a square matrix module (\texttt{SQUAREMATRIX})
  and returns a module that implements \texttt{TTTSTATE}. The functors in
  \texttt{tictactoe.sml} effectively take a module structure that implements the
  previous module of the game pattern and produce the current (or next)
  modular piece.

  The functor \texttt{TttActionFn} (used by modules structures like
  \texttt{TttAction} and \texttt{Ttt3DAction}) takes a module that implements
  \texttt{TTTSTATE} and gives us a module implementing \texttt{TTTACTION}.
  The functors \texttt{TttRandomAgent} and \texttt{TttHumanAgentFn} take
  modules which implement \texttt{TTTACTION} and return a module that implements
  \texttt{AGENT}.

  We can then instantiate a structure (like \texttt{TttExecRandom} for example)
  that will run a game of \textit{Tic-tac-toe}
  To do so we provide our \texttt{ExecFn} functor from
  \texttt{game.sml} with a module that implements \texttt{AGENT}.
  \texttt{ExecFn} then gives us a module that implements \texttt{EXEC}.

@@

  \subsection{You can do it in 2-dimensions, but can you do it in n-dimensions!}
  Initially, we had conceived on implementing a pokemon battle simulator. As we
  dove into the problem domain we realized that instead of falling into a
  specific, flow-chart design with modules like: Pokemon, Combat, Moves etc. it
  would be better design to create an abstract game engine and then make pokemon
  \textit{an instance of} the abstract game engine. A similar pattern occurred
  during our first implementation of the matrix library utilized by the tic tac
  toe state viz. we realized we wrote a, too specific, flowchart-y instance of a
  matrix library. Thus, in the spirit of modularity, we refactored the matrix
  library to implement an abstract \texttt{CONTAINER} signature. Then based on
  this signature, it became possible to create sub-signatures that specified the
  types of the CONTAINER signature. A good example is shown below:

@@

  \end{verbatim}
  In this example we show the relevant parts of the \texttt{CONTAINER} signature
  and how we used that signature to create a \textit{more specific}
  \texttt{SQUAREMATRIX} signature e.g. a matrix whose size datatype need only be
  a single integer value. This allows for a very flexible and modular
  implementation, in addition to square matrices we could create jagged matrices
  by following the same pattern viz. defining a \texttt{JAGGEDMATRIX} and
  setting size to ``int * int''.

>>>>

  As we touched on before, the notion of a game is generalized in the
  \texttt{game.sml} file. This defines the signatures which make up the pattern
  of a game and a functor to run a game.
  The general pattern for a game under our model consists of three purposely
  isolated pieces.
  These are \textit{State}, \textit{Actions}, and \textit{Agents} - which are
  each given their own signature in \texttt{game.sml}.
  These are then wrapped together with a functor to execute a game.
  The idea is that any game consists of a state, a set of actions on that state,
  and agents that select which action to perform on a given state. A specific implementation of a
  game would use the relationship between these three general pieces to define
  a runnable game with its own modules.

@@

  In \texttt{tictactoe.sml} we have two signatures: one that extends state
  (the \texttt{STATE} signature from \texttt{game.sml}) for \textit{Tic-tac-toe}
  called \texttt{TTTSTATE} and another that does the same with action
  (the \texttt{ACTION} signature from \texttt{game.sml}) called
  \texttt{TTTACTION}.
  We don't extend the \texttt{AGENT} signature specifically for \textit{Tic-tac-toe}.
  These extended signatures are perfectly compatible with the original signatures
  due to the structural matching of SML signatures.

  There are many module structures instantiated in \texttt{tictactoe.sml}
  for the various kinds of state, actions and agents that we eventually want to
  use---such as \texttt{TttState} and \texttt{TttAction} for instance---but each will
  be used to implement the next kind of module.
  These module structures are bound to functors like \texttt{TttStateFn} which
  takes a module implementing a square matrix module (\texttt{SQUAREMATRIX})
  and returns a module that implements \texttt{TTTSTATE}. The functors in
  \texttt{tictactoe.sml} effectively take a module structure that implements the
  previous module of the game pattern and produce the current (or next)
  modular piece.

  The functor \texttt{TttActionFn} (used by module structures like
  \texttt{TttAction} and \texttt{Ttt3DAction}) takes a module that implements
  \texttt{TTTSTATE} and gives us a module implementing \texttt{TTTACTION}.
  The functors \texttt{TttRandomAgentFn} and \texttt{TttHumanAgentFn} take
  modules which implement \texttt{TTTACTION} and return a module that implements
  \texttt{AGENT}.

  We can then instantiate a structure (like \texttt{TttExecRandom} for example)
  that will run a game of \textit{Tic-tac-toe}.
  To do so we provide our \texttt{ExecFn} functor from
  \texttt{game.sml} with a module that implements \texttt{AGENT}.
  \texttt{ExecFn} then gives us a module that implements \texttt{EXEC}.

@@

  \subsection{You can do it in 2-dimensions, but can you do it in n-dimensions!}
  Our initial intent was to implement a pokemon battle simulator. As we
  dove into the problem domain we realized that instead of falling into a
  specific, flow-chart design with modules like: Pokemon, Combat, Moves etc. it
  would be better design to create an abstract game engine and then make pokemon
  \textit{an instance of} the abstract game engine. A similar pattern occurred
  during our first implementation of the matrix library utilized by the tic tac
  toe state. We wrote a single matrix module with more rigidly defined 
  dimensions and underlying implementation details. While it was abstract, it was not very modular,
  as it required writing a new module from the ground up if any of the 
  implementation details were to be changed. Thus, in the spirit of modularity, we refactored the matrix
  library to implement an abstract \texttt{CONTAINER} signature. Then based on
  this signature, it became possible to create sub-signatures that specified the
  types of the CONTAINER signature. A good example is shown below:

@@

  \end{verbatim}
  In this example we show the relevant parts of the \texttt{CONTAINER} signature
  and how we used that signature to create a \textit{more specific}
  \texttt{SQUAREMATRIX} signature e.g. a matrix whose size datatype need only be
  a single integer value. This allows for a very flexible and modular
  implementation, in addition to square matrices we could create jagged two-dimensional matrices
  by following the same pattern viz. defining a \texttt{JAGGEDMATRIX} and
  setting size to ``int * int''.
